---
title: "Lesson 12: Interactions, Part 2"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#213c96"
date: "`r library(here); source(here('class_dates.R')); w8d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lesson 12: Interactions 2"
    highlight-style: ayu
    html-math-method: mathjax
execute:
  freeze: auto  # re-render only when source changes
---

```{r}
#| label: "setup" 
#| include: false
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(knitr)
library(broom)
library(rstatix)
library(gt)
library(readxl)
#----------
# new packages
# install.packages("describedata")
library(describedata) # gladder()
library(gridExtra)   # grid.arrange()
library(ggfortify)  # autoplot(model)
# New Day 6
library(gtsummary)

# New Day 7
library(plotly) # for plot_ly() command
library(GGally) # for ggpairs() command 
library(ggiraphExtra)   # for ggPredict() command
library(here)

# Load the data - update code if the csv file is not in the same location on your computer
# If you need to download the file, please go to ur shared folder under Data > Slides
load(here("data", "gapminder2022.RData"))
gapm = gapm %>%
  mutate(fs_order = case_when(freedom_status == "F" ~ 3,
                                freedom_status == "PF" ~ 2,
                                freedom_status == "NF" ~ 1), 
         freedom_status = fct_relevel(freedom_status, "NF", "PF", "F")) %>%
  mutate(income_level_4 = fct_relevel(income_level_4, 
                                      "Low income", 
                                      "Lower middle income",
                                      "Upper middle income",
                                      "High income")) %>%
  mutate(income_level_2 = case_when(income_level_4 %in% c("Low income", "Lower middle income") ~ "Lower income",
                                      income_level_4 %in% c("Upper middle income", "High income") ~ "Higher income")) %>%
  mutate(income_level_2 = fct_relevel(income_level_2, "Lower income", "Higher income")) %>%
  mutate(CP_c = cell_phones_100 - mean(cell_phones_100, na.rm=T))

mlr1 <- gapm %>% lm(formula = life_exp ~ cell_phones_100 + vax_rate)

gapm2 = gapm %>% select(cell_phones_100, life_exp, vax_rate, freedom_status, income_level_4, basic_sani)
```


# Learning Objectives

**Last time:**

1.  Define confounders and effect modifiers, and how they interact with the main relationship we model.
2.  Interpret the interaction component of a model with **a binary categorical covariate and continuous covariate**, and how the main variable's effect changes.
3.  Interpret the interaction component of a model with **a multi-level categorical covariate and continuous covariate**, and how the main variable's effect changes.

**This time:**

4.  Interpret the interaction component of a model with **two categorical covariates**, and how the main variable's effect changes.
5.  Interpret the interaction component of a model with **two continuous covariates**, and how the main variable's effect changes.
6. Report results for a best-fit line (with confidence intervals) at different levels of an effect measure modifier

# Learning Objectives

**Last time:**

1.  Define confounders and effect modifiers, and how they interact with the main relationship we model.
2.  Interpret the interaction component of a model with **a binary categorical covariate and continuous covariate**, and how the main variable's effect changes.
3.  Interpret the interaction component of a model with **a multi-level categorical covariate and continuous covariate**, and how the main variable's effect changes.

**This time:**

::: lob
4.  Interpret the interaction component of a model with **two categorical covariates**, and how the main variable's effect changes.
:::

5.  Interpret the interaction component of a model with **two continuous covariates**, and how the main variable's effect changes.
6. Report results for a best-fit line (with confidence intervals) at different levels of an effect measure modifier

## Do we think income level can be an effect modifier for freedom status?

::: columns
::: column
-   Taking a break from cell phones to demonstrate interactions for two categorical variables

-   We can start by visualizing the relationship between life expectancy and freedom status *by income level*

-   Questions of interest: Does the effect of freedom status on life expectancy differ depending on income level?

    -   This is the same as: Is income level an effect modifier for freedom status?

-   Let's run an interaction model to see!
:::

::: column
```{r fig.height=7, fig.width=9, warning=F, fig.align='center'}
#| echo: false
ggplot(gapm, aes(x = freedom_status, y = life_exp, 
                  color = income_level_2)) +
  geom_jitter(size = 2, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  labs(x = "Freedom Status", 
       y = "Life expectancy (years)",
       title = "Life expectancy vs. Freedom Status", 
       color = "Income level") +
  theme(axis.title = element_text(size = 20), 
        axis.text = element_text(size = 20), 
        title = element_text(size = 20), 
        legend.text=element_text(size= 18)) +
  scale_color_manual(values=c("#FF8021", "#34AC8B"))
```
:::
:::

## Model with interaction between a *multi-level categorical and binary variables*

Model we are fitting:

$$\begin{aligned}LE = &\beta_0 + \beta_1 I(\text{high income}) + \beta_2 I(\text{FS} = \text{PF}) + \beta_3 I(\text{FS} = \text{F})  + \\ & \beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \beta_5 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{F}) + \epsilon \end{aligned}$$

-   $LE$ as life expectancy
-   $I(\text{high income})$ as indicator of high income
-   $I(\text{FS} = \text{PF})$ and $I(\text{FS} = \text{F})$ as the indicator for each freedom status

In R:

```{r}
#| echo: true
m_int_wr_inc = gapm %>% 
  lm(formula = life_exp ~ income_level_2 + freedom_status + income_level_2*freedom_status)
m_int_wr_inc = gapm %>% 
  lm(formula = life_exp ~ income_level_2*freedom_status)
```

## Displaying the regression table and writing fitted regression equation

```{r}
#| echo: true

tidy_m_int_wr_inc = tidy(m_int_wr_inc, conf.int=T)
tidy_m_int_wr_inc %>% gt() %>% tab_options(table.font.size = 35) %>% fmt_number(decimals = 3)
```

$$\begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & `r round(tidy_m_int_wr_inc$estimate[1], 2)` + `r round(tidy_m_int_wr_inc$estimate[2], 2)` \cdot I(\text{high income}) + `r round(tidy_m_int_wr_inc$estimate[3], 2)` \cdot I(\text{FS} = \text{PF})  `r round(tidy_m_int_wr_inc$estimate[4], 2)` \cdot I(\text{FS} = \text{F}) + \\ & `r round(tidy_m_int_wr_inc$estimate[5], 2)` \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + `r round(tidy_m_int_wr_inc$estimate[6], 2)`\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})\\
\end{aligned}$$

## Poll Everywhere Question 4

::: columns
::: {.column width="60%"}
:::

::: {.column width="40%"}
```{r fig.height=8, fig.width=8, warning=F, fig.align='center'}
#| echo: false
ggplot(gapm, aes(x = freedom_status, y = life_exp, 
                  color = income_level_2)) +
  geom_jitter(size = 2, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  labs(x = "Freedom Status", 
       y = "Life expectancy (years)",
       title = "Life expectancy vs. Freedom Status", 
       color = "Income level") +
  theme(axis.title = element_text(size = 20), 
        axis.text = element_text(size = 20), 
        title = element_text(size = 20), 
        legend.text=element_text(size= 18), legend.position = "bottom") +
  scale_color_manual(values=c("#FF8021", "#34AC8B"))
```
:::
:::


## Comparing fitted regression *means* for each freedom status

$$\begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & `r round(tidy_m_int_wr_inc$estimate[1], 2)` + `r round(tidy_m_int_wr_inc$estimate[2], 2)` \cdot I(\text{high income}) + `r round(tidy_m_int_wr_inc$estimate[3], 2)` \cdot I(\text{FS} = \text{PF})  `r round(tidy_m_int_wr_inc$estimate[4], 2)` \cdot I(\text{FS} = \text{F}) + \\ & `r round(tidy_m_int_wr_inc$estimate[5], 2)` \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + `r round(tidy_m_int_wr_inc$estimate[6], 2)`\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})
\end{aligned}$$

::::::::::::::::::: columns
:::::: {.column width="33%"}
::::: proof1
::: proof-title
Not free
:::

::: proof-cont
$$\begin{aligned} 
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \\ 
& \widehat\beta_2 \cdot 0 + \widehat\beta_3 \cdot 0  + \\ 
& \widehat\beta_4 \cdot I(\text{high income}) \cdot 0 + \\
& \widehat\beta_5\cdot I(\text{high income}) \cdot 0 \\ 
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income})
\end{aligned}$$
:::
:::::
::::::

:::::: {.column width="33%"}
::::: fact
::: fact-title
Partly free
:::

::: fact-cont
$$\begin{aligned} 
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \\ 
& \widehat\beta_2 \cdot 1 + \widehat\beta_3 \cdot 0  + \\ 
& \widehat\beta_4 \cdot I(\text{high income}) \cdot 1 + \\
& \widehat\beta_5\cdot I(\text{high income}) \cdot 0 \\ 
\widehat{LE} = & \big(\widehat\beta_0 + \widehat\beta_2 \big) + \\ & \big(\widehat\beta_1 + \widehat\beta_4  \big) I(\text{high income}) \\
\end{aligned}$$
:::
:::::
::::::


:::::: {.column width="33%"}
::::: definition
::: def-title
Free
:::

::: def-cont
$$\begin{aligned} 
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \\ 
& \widehat\beta_2 \cdot 1 + \widehat\beta_3 \cdot 0  + \\ 
& \widehat\beta_4 \cdot I(\text{high income}) \cdot 1 + \\
& \widehat\beta_5\cdot I(\text{high income}) \cdot 0 \\ 
\widehat{LE} = & \big(\widehat\beta_0 + \widehat\beta_3 \big) + \\ & \big(\widehat\beta_1 + \widehat\beta_5  \big) I(\text{high income}) \\
\end{aligned}$$
:::
:::::
::::::
:::::::::::::::::::
## Comparing fitted regression *means* for each income level {.smaller}

$$\begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 I(\text{high income}) + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & `r round(tidy_m_int_wr_inc$estimate[1], 2)` + `r round(tidy_m_int_wr_inc$estimate[2], 2)` \cdot I(\text{high income}) + `r round(tidy_m_int_wr_inc$estimate[3], 2)` \cdot I(\text{FS} = \text{PF})  `r round(tidy_m_int_wr_inc$estimate[4], 2)` \cdot I(\text{FS} = \text{F}) + \\ & `r round(tidy_m_int_wr_inc$estimate[5], 2)` \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + `r round(tidy_m_int_wr_inc$estimate[6], 2)`\cdot I(\text{high income}) \cdot I(\text{FS} = \text{F})
\end{aligned}$$

::: columns
::: column
::: proof1
::: proof-title
For lower income countries: $I(\text{high income})=0$
:::

::: proof-cont
$$ \begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 \cdot 0 + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot 0 \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5 \cdot 0 \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  \\
\end{aligned}$$
:::
:::
:::

::: column
::: fact
::: fact-title
For higher income countries: $I(\text{high income}) =1$
:::

::: fact-cont
$$ \begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 \cdot 1 + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot 1 \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5 \cdot 1 \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & \big(\widehat\beta_0 + \widehat\beta_1 \big) + \big(\widehat\beta_2 + \widehat\beta_4 \big) I(\text{FS} = \text{PF}) + \\ & \big(\widehat\beta_3 + \widehat\beta_5 \big) I(\text{FS} = \text{F})  \\
\end{aligned}$$
:::
:::
:::
:::

- Example interpretation: The America's effect on mean life expectancy increases $\widehat{\beta}_5$ comparing high income to low income countries.

## Let's take a look back at the plot

::: columns
::: {.column width="50%"}
::: proof1
::: proof-title
For lower income countries: $I(\text{high income}) =0$
:::

::: proof-cont
$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F}) \\
\end{aligned}$$
:::
:::

::: fact
::: fact-title
For higher income countries: $I(\text{high income}) =1$
:::

::: fact-cont
$$ \begin{aligned}
\widehat{LE} = & \big(\widehat\beta_0 + \widehat\beta_1 \big) + \big(\widehat\beta_2 + \widehat\beta_4 \big) I(\text{FS} = \text{PF}) + \\ & \big(\widehat\beta_3 + \widehat\beta_5 \big) I(\text{FS} = \text{F}) \\
\end{aligned}$$
:::
:::
:::

::: {.column width="50%"}
```{r fig.height=8.5, fig.width=8.5, warning=F, fig.align='center'}
#| echo: false
ggplot(gapm, aes(x = freedom_status, y = life_exp, 
                  color = income_level_2)) +
  geom_jitter(size = 2, alpha = .6, width = 0.2) +
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18) +
  labs(x = "Freedom Status", 
       y = "Life expectancy (years)",
       title = "Life expectancy vs. Freedom Status", 
       color = "Income level") +
  theme(axis.title = element_text(size = 20), 
        axis.text = element_text(size = 20), 
        title = element_text(size = 20), 
        legend.text=element_text(size= 18), legend.position = "bottom") +
  scale_color_manual(values=c("#FF8021", "#34AC8B"))
```
:::
:::

## Interpretation for interaction between two categorical variables

$$ \begin{aligned}
\widehat{LE} = &\widehat\beta_0 + \widehat\beta_1 \cdot 1 + \widehat\beta_2 I(\text{FS} = \text{PF}) + \widehat\beta_3 I(\text{FS} = \text{F})  + \\ & \widehat\beta_4 \cdot 1 \cdot I(\text{FS} = \text{PF}) + \widehat\beta_5 \cdot 1 \cdot I(\text{FS} = \text{F})\\
\widehat{LE} = & \bigg[\widehat\beta_0 + \widehat\beta_1 \cdot I(\text{high income})\bigg]  + \bigg[\widehat\beta_2 + \widehat\beta_4 \cdot I(\text{high income})\bigg] I(\text{FS} = \text{PF}) + \\ & \bigg[\widehat\beta_3 + \widehat\beta_5 \cdot I(\text{high income})\bigg]  I(\text{FS} = \text{F}) \\
\end{aligned}$$

-   Interpretation:
    -   $\widehat\beta_1$ = mean change in life expectancy comparing high income to low income countries, for countries/territories that are [**not free**]{style="color: #FF8021"}
    -   $\widehat\beta_4$ = mean change in life expectancy comparing high income to low income countries, for countries/territories that are [**partly free**]{style="color: #34AC8B"}
    -   $\widehat\beta_5$ = mean change in life expectancy comparing high income to low income countries, for countries/territories that are [**free**]{style="color: #4FADF3"}

## Test interaction between two categorical variables (1/2)

-   We run an F-test for a group of coefficients ($\beta_4$, $\beta_5$) in the below model (see lesson 9)

$$\begin{aligned}LE = &\beta_0 + \beta_1 I(\text{high income}) + \beta_2 I(\text{FS} = \text{PF}) + \beta_3 I(\text{FS} = \text{F})  + \\ & \beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \beta_5 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{F}) + \epsilon \end{aligned}$$

::: columns
::: {.column width="42%"}
::: proof1
::: proof-title
Null $H_0$
:::

::: proof-cont
$\beta_4= \beta_5 =0$
:::
:::
:::

::: {.column width="58%"}
::: definition
::: def-title
Alternative $H_1$
:::

::: def-cont
$\beta_4\neq0$ and/or $\beta_5\neq0$ 
:::
:::
:::
:::

::: columns
::: {.column width="42%"}
::: proof1
::: proof-title
Null / Smaller / Reduced model
:::

::: proof-cont
$$\begin{aligned}LE = &\beta_0 + \beta_1 I(\text{high income}) + \\
&\beta_2 I(\text{FS} = \text{PF}) + \beta_3 I(\text{FS} = \text{F}) + \\
& \epsilon \end{aligned}$$
:::
:::
:::

::: {.column width="58%"}
::: definition
::: def-title
Alternative / Larger / Full model
:::

::: def-cont
$$\begin{aligned}LE = &\beta_0 + \beta_1 I(\text{high income}) + \beta_2 I(\text{FS} = \text{PF}) + \\ & \beta_3 I(\text{FS} = \text{F})  + \beta_4 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{PF}) + \\ & \beta_5 \cdot I(\text{high income}) \cdot I(\text{FS} = \text{F}) + \epsilon \end{aligned}$$
:::
:::
:::
:::

## Test interaction between two categorical variables (2/2)

-   Fit the reduced and full model

```{r}
#| echo: true
m_int_wr_inc_red = gapm %>% 
  lm(formula = life_exp ~ income_level_2 + freedom_status)
m_int_wr_inc_full = gapm %>% 
  lm(formula = life_exp ~ income_level_2*freedom_status)
```

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Display the ANOVA table with F-statistic and p-value"

anova_table = anova(m_int_wr_inc_red, m_int_wr_inc_full)
anova_table %>% tidy() %>% gt() %>% tab_options(table.font.size = 35) %>% 
  fmt_number(decimals = 3)
```

-   **Conclusion:** There is not a significant interaction between freedom status and income level (p = `r round(anova_table[2,6], 3)`). 

# Learning Objectives

**Last time:**

1.  Define confounders and effect modifiers, and how they interact with the main relationship we model.
2.  Interpret the interaction component of a model with **a binary categorical covariate and continuous covariate**, and how the main variable's effect changes.
3.  Interpret the interaction component of a model with **a multi-level categorical covariate and continuous covariate**, and how the main variable's effect changes.

**This time:**

4.  Interpret the interaction component of a model with **two categorical covariates**, and how the main variable's effect changes.

::: lob
5.  Interpret the interaction component of a model with **two continuous covariates**, and how the main variable's effect changes.
:::

6. Report results for a best-fit line (with confidence intervals) at different levels of an effect measure modifier

## Do we think vaccination rate is an effect modifier for cell phones?

::: columns
::: {.column width="40%"}
-   We can start by visualizing the relationship between life expectancy and cell phones *by vaccination rate*

-   Questions of interest: Does the effect of cell phones on life expectancy differ depending on vaccination rate?

    -   This is the same as: Is vaccination rate is an effect modifier for cell phones? Is vaccination rate an effect modifier of the association between life expectancy and cell phones?

-   Let's run an interaction model to see!
:::

::: {.column width="60%"}
```{r fig.height=4, fig.width=6, warning=F, fig.align='center'}
#| echo: false
m_int_vr = lm(life_exp ~ cell_phones_100*vax_rate, data = gapm)
(mr1_2d = ggPredict(m_int_vr, interactive = T) )
```
:::
:::

## Model with interaction between *two continuous variables*

Model we are fitting:

$$ LE = \beta_0 + \beta_1 CP^c + \beta_2 VR^c + \beta_3 CP^c \cdot VR^c + \epsilon$$

-   $LE$ as life expectancy
-   $CP^c$ as the **centered** around the mean number of cell phones per 100 people (continuous variable)
-   $VR^c$ as the **centered** around the mean vaccination rate (continuous variable)

```{r}
#| code-fold: true
#| code-summary: "Code to center CP and VR"
#| echo: true
gapm = gapm %>% 
  mutate(CP_c = cell_phones_100 - mean(cell_phones_100), 
         VR_c = vax_rate - mean(vax_rate))
mean_VR = mean(gapm$vax_rate) %>% round(digits = 2)
mean_CP = mean(gapm$cell_phones_100) %>% round(digits = 2)
```

In R:

```{r}
#| echo: true

m_int_vr = gapm %>% lm(formula = life_exp ~ CP_c + VR_c + CP_c*VR_c)
```

OR

```{r}
#| echo: true

m_int_vr = gapm %>% lm(formula = life_exp ~ CP_c*VR_c)
```

## Displaying the regression table and writing fitted regression equation

```{r}
#| echo: true
tidy_m_vr = tidy(m_int_vr, conf.int=T) 
tidy_m_vr %>% gt() %>% tab_options(table.font.size = 35) %>% fmt_number(decimals = 5)
```

$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \widehat\beta_2 VR^c + \widehat\beta_3 CP^c \cdot VR^c \\
\widehat{LE} = & `r round(tidy_m_vr$estimate[1],2)` + `r round(tidy_m_vr$estimate[2],2)` \cdot CP^c + `r round(tidy_m_vr$estimate[3],2)` \cdot VR^c + `r round(tidy_m_vr$estimate[4],3)` \cdot CP^c \cdot VR^c
\end{aligned}$$

## Comparing fitted regression lines for various vaccination rates

$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \widehat\beta_2 VR^c + \widehat\beta_3 CP^c \cdot VR^c \\
\widehat{LE} = & `r round(tidy_m_vr$estimate[1],2)` + `r round(tidy_m_vr$estimate[2],2)` \cdot CP^c + `r round(tidy_m_vr$estimate[3],2)` \cdot VR^c + `r round(tidy_m_vr$estimate[4],3)` \cdot CP^c \cdot VR^c
\end{aligned}$$

To identify different lines, we need to pick example vaccination rates:

```{r}
#| echo: false

vax_diff = 5
vax_rate1 = mean_VR - vax_diff
vax_rate2 = mean_VR
vax_rate3 = mean_VR + vax_diff
```

::: columns
::: {.column width="33%"}
::: FS1
::: FS1-title
Vaccination rate of `r vax_rate1` \%
:::

::: FS1-cont
$$\begin{aligned} 
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \\ & \widehat\beta_2 \cdot (-`r vax_diff`) + \\ & \widehat\beta_3 CP^c \cdot (-`r vax_diff`) \\
\widehat{LE} = & \big(\widehat\beta_0 - `r vax_diff` \widehat\beta_2 \big)+ \\ & \big(\widehat\beta_1 - `r vax_diff` \widehat\beta_3  \big) CP^c
\end{aligned}$$
:::
:::
:::

::: {.column width="33%"}
::: FS2
::: FS2-title
Vaccination rate of `r mean_VR` \%
:::

::: FS2-cont
$$\begin{aligned} 
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \\ & \widehat\beta_2 \cdot 0 + \\ & \widehat\beta_3 CP^c \cdot 0 \\
\widehat{LE} = & \big(\widehat\beta_0 \big)+ \\ & \big(\widehat\beta_1 \big) CP^c
\end{aligned}$$
:::
:::
:::

::: {.column width="33%"}
::: FS3
::: FS3-title
Vaccination rate of `r vax_rate3` \%
:::

::: FS3-cont
$$\begin{aligned} 
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \\ & \widehat\beta_2 \cdot `r vax_diff` + \\ & \widehat\beta_3 CP^c \cdot `r vax_diff` \\
\widehat{LE} = & \big(\widehat\beta_0 + `r vax_diff` \widehat\beta_2 \big)+ \\ & \big(\widehat\beta_1 + `r vax_diff` \widehat\beta_3  \big) CP^c
\end{aligned}$$
:::
:::
:::
:::

## Poll Everywhere Question??

## Interpretation for interaction between two continuous variables

$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \widehat\beta_2 VR^c + \widehat\beta_3 CP^c \cdot VR^c \\
\widehat{LE} = & \bigg[\widehat\beta_0 + \widehat\beta_2 \cdot VR^c \bigg] + \underbrace{\bigg[\widehat\beta_1 + \widehat\beta_3 \cdot VR^c \bigg]}_\text{CP's effect} CP \\
\end{aligned}$$

-   Interpretation:

    -   $\beta_3$ = mean change in cell phones's effect, for every 1 \% increase in vaccination rate

-   In summary, the interaction term can be interpreted as "difference in adjusted effect of number of cell phones for every 1 \% increase in vaccination rate"

-   It will be helpful to test the interaction to round out this interpretation!!

## Test interaction between two continuous variables (1/2)

-   We run an F-test for a single coefficients ($\beta_3$) in the below model (see lesson 9)

$$ LE = \beta_0 + \beta_1 CP^c + \beta_2 VR^c + \beta_3 CP^c \cdot VR^c + \epsilon$$

::: columns
::: {.column width="10%"}
:::

::: {.column width="40%"}
::: proof1
::: proof-title
Null $H_0$
:::

::: proof-cont
$$\beta_3=0$$
:::
:::
:::

::: {.column width="40%"}
::: definition
::: def-title
Alternative $H_1$
:::

::: def-cont
$$\beta_3\neq0$$
:::
:::
:::

::: {.column width="10%"}
:::
:::

::: columns
::: {.column width="10%"}
:::

::: {.column width="40%"}
::: proof1
::: proof-title
Null / Smaller / Reduced model
:::

::: proof-cont
$$ LE = \beta_0 + \beta_1 CP^c + \beta_2 VR^c + \epsilon$$
:::
:::
:::

::: {.column width="40%"}
::: definition
::: def-title
Alternative / Larger / Full model
:::

::: def-cont
$$\begin{aligned}
LE = & \beta_0 + \beta_1 CP^c + \beta_2 VR^c + \\ & \beta_3 CP^c \cdot VR^c + \epsilon
\end{aligned}$$
:::
:::
:::
:::

## Test interaction between two continuous variables (2/2)

-   Fit the reduced and full model

```{r}
#| echo: true
m_int_vr_red = gapm %>% 
  lm(formula = life_exp ~ CP_c + VR_c)
m_int_vr_full = gapm %>%
  lm(formula = life_exp ~ CP_c + VR_c + CP_c*VR_c)
```

 

```{r}
#| code-fold: true
#| code-summary: "Display the ANOVA table with F-statistic and p-value"
#| echo: true

anova_table2 = anova(m_int_vr_red, m_int_vr_full) 
anova_table2 %>% tidy() %>% gt() %>% tab_options(table.font.size = 35) %>% 
  fmt_number(decimals = 3)
```

-   Conclusion: There is not a significant interaction between cell phones and vaccination rate (p = `r round(anova_table2[2,6], 3)`). Vaccination rate is not an effect modifier of the association between cell phones and life expectancy.

# Learning Objective

**Last time:**

1.  Define confounders and effect modifiers, and how they interact with the main relationship we model.
2.  Interpret the interaction component of a model with **a binary categorical covariate and continuous covariate**, and how the main variable's effect changes.
3.  Interpret the interaction component of a model with **a multi-level categorical covariate and continuous covariate**, and how the main variable's effect changes.

**This time:**

4.  Interpret the interaction component of a model with **two categorical covariates**, and how the main variable's effect changes.
5.  Interpret the interaction component of a model with **two continuous covariates**, and how the main variable's effect changes.

::: lob
6. Report results for a best-fit line (with confidence intervals) at different levels of an effect measure modifier
:::

## How to find the confidence interval for each slope?

- In the example with VR and CP, we showed:

::: FS3
::: FS3-title
Best-fit line for vaccination rate of `r vax_rate3` \%
:::
::: FS3-cont
$$\begin{aligned} 
\widehat{LE} = & \big(\widehat\beta_0 + `r vax_diff` \widehat\beta_2 \big)+  \big(\widehat\beta_1 + `r vax_diff` \widehat\beta_3  \big) CP^c
\end{aligned}$$
:::
:::

- Often, we want to report the estimate of the combined coefficients: $\widehat\beta_1 + `r vax_diff` \widehat\beta_3$

  - This allows us to make a statement like: "At a vaccination rate of `r vax_rate3`\%, mean life expectancy increases $\big(\widehat\beta_1 + `r vax_diff` \widehat\beta_3 \big)$ years for every one additional cell phone per 100 people (95% CI: \_\_, \_\_)."

 

- We can calculate $\widehat\beta_1 + `r vax_diff` \widehat\beta_3$ by using the values of the estimated coefficients

- BUT we always want to have a **95% confidence interval** when we report this combined estimate!!

## Getting a 95% confidence interval requires linear combinations!

- If we want a confidence interval for $\widehat\beta_1 + `r vax_diff` \widehat\beta_3$, then we would use the formula:

$$\bigg(\widehat\beta_1 + `r vax_diff` \widehat\beta_3 \bigg) \pm t^* \times SE_{(\beta_1 + `r vax_diff` \beta_3)}$$

- The hard part is figuring out what $SE_{(\beta_1 + `r vax_diff` \beta_3)}$ (or $\text{Var}(\beta_1 + `r vax_diff` \beta_3)$) equals

- We need to go back to variance of linear combinations (BSTA 511/611, EPI 525):
$$\text{Var}(aX + bY) = a^2\text{Var}(X) + b^2\text{Var}(Y) + 2ab\text{Cov}(X, Y)$$
or
$$\text{Var}(aX - bY) = a^2\text{Var}(X) + b^2\text{Var}(Y) - 2ab\text{Cov}(X, Y)$$

## Calculating $SE_{(\beta_1 + `r vax_diff` \beta_3)}$ "by hand" (REFERENCE)

- A helpful function that returns the variance-covariance matric of all the coefficients in model `m_int_vr`:

::: columns
::: {.column width="70%"}
```{r}
#| echo: true
vcov(m_int_vr)
```
:::
::: {.column width="30%"}
$$ \begin{aligned}
\text{Var}(\beta_1) & = `r round(vcov(m_int_vr)[2,2], 9)` \\
\text{Var}(\beta_3) & = `r round(vcov(m_int_vr)[3,3], 9)` \\
\text{Cov}(\beta_1, \beta_3) & = `r round(vcov(m_int_vr)[2,3], 9)` \\
\end{aligned}$$
:::
:::

$$ \begin{aligned}
\text{Var}(\beta_1 + `r vax_diff` \beta_3) & = \text{Var}(\beta_1) + `r vax_diff` ^2 \text{Var}(\beta_3) + 2 \cdot `r vax_diff` \cdot \text{Cov}(\beta_1, \beta_3) \\ 
\text{Var}(\beta_1 + `r vax_diff` \beta_3) & = `r round(vcov(m_int_vr)[2,2], 9)` + `r vax_diff`^2 \ \times `r round(vcov(m_int_vr)[4,4], 9)` + 2 \cdot `r vax_diff` \cdot `r round(vcov(m_int_vr)[2,4], 9)` \\ 
\text{Var}(\beta_1 + `r vax_diff` \beta_3) & = `r round(vcov(m_int_vr)[2,2] + vax_diff^2 * vcov(m_int_vr)[4,4] + 2 * vax_diff * vcov(m_int_vr)[2,4], 9)` \\
SE_{(\beta_1 + `r vax_diff` \beta_3)} & = \sqrt{`r round(vcov(m_int_vr)[2,2] + vax_diff^2 * vcov(m_int_vr)[4,4] + 2 * vax_diff * vcov(m_int_vr)[2,4], 9)`} \\ 
SE_{(\beta_1 + `r vax_diff` \beta_3)} & = `r round(sqrt(vcov(m_int_vr)[2,2] + vax_diff^2 * vcov(m_int_vr)[4,4] + 2 * vax_diff * vcov(m_int_vr)[2,4]), 9)`
\end{aligned}$$


## We can use R and `estimable()` to find the estimate and CI

For $\widehat\beta_1 + `r vax_diff` \widehat\beta_3$:

```{r}
#| echo: true

library(gmodels)
m_int_vr %>% estimable(
                   c("(Intercept)" = 0,      # beta0
                     "CP_c"       = 1,      # beta1
                     "VR_c"        = 0,      # beta2
                     "CP_c:VR_c"  = 5),  # beta3
                   conf.int = 0.95)
```

```{r}
m_int_vr_101 = m_int_vr %>% estimable(
                   c("(Intercept)" = 0,      # beta0
                     "CP_c"       = 1,      # beta1
                     "VR_c"        = 0,      # beta2
                     "CP_c:VR_c"  = 5),  # beta3
                   conf.int = 0.95)
```

 

**Our conclusion:** At a vaccination rate of `r vax_rate3` \%, mean life expectancy increases `r round(m_int_vr_101$Estimate, 3)` years for every one additional cell phone per 100 people (95% CI: `r round(m_int_vr_101$Lower.CI, 3)`, `r round(m_int_vr_101$Upper.CI, 3)`).

## Another example: income (binary) and CP (1/2)

```{r}
#| echo: true

m_int_inc2 = gapm %>% 
  lm(formula = life_exp ~ CP_c*income_level_2)
```

```{r}
coef_centered = tidy(m_int_inc2, conf.int=T)
```

$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP^c + \widehat\beta_2 I(\text{high income}) + \widehat\beta_3 CP^c \cdot I(\text{high income}) \\
\widehat{LE} = & `r round(coef_centered$estimate[1], 3)` + `r round(coef_centered$estimate[2], 3)` \cdot CP^c + `r round(coef_centered$estimate[3], 3)` \cdot I(\text{high income}) `r round(coef_centered$estimate[4], 3)` \cdot CP^c \cdot I(\text{high income})
\end{aligned}$$

::: columns
::: column
::: proof1
::: proof-title
For lower income countries: $I(\text{high income}) =0$
:::

::: proof-cont
$$ \begin{aligned}
\widehat{LE} = & \widehat\beta_0 + \widehat\beta_1 CP\\
\widehat{LE} = & `r round(coef_centered$estimate[1], 2)` + `r round(coef_centered$estimate[2], 3)` \cdot CP
\end{aligned}$$
:::
:::
:::

::: column
::: fact
::: fact-title
For higher income countries: $I(\text{high income}) =1$
:::

::: fact-cont
$$ \begin{aligned}
\widehat{LE} = & (\widehat\beta_0 +\widehat\beta_2) + (\widehat\beta_1 +\widehat\beta_3) CP \\
\widehat{LE} = & (`r round(coef_centered$estimate[1], 2)` + `r round(coef_centered$estimate[3], 3)`) + (`r round(coef_centered$estimate[2], 3)` `r round(coef_centered$estimate[4], 3)`) \cdot CP \\
\widehat{LE} = & `r round(coef_centered$estimate[1] + coef_centered$estimate[3], 2)` + `r round(coef_centered$estimate[2] +coef_centered$estimate[4], 3)` \cdot CP
\end{aligned}$$
:::
:::
:::
:::

## Another example: income (binary) and CP (2/2)


```{r}
#| echo: true

names(m_int_inc2$coefficients) # I just need to see the exact names
m_int_inc2 %>% estimable(
                   c("(Intercept)" = 0,      # beta0
                     "CP_c"       = 1,      # beta1
                     "income_level_2Higher income" = 0,      # beta2
                     "CP_c:income_level_2Higher income"  = 1),  # beta3
                   conf.int = 0.95)
```

```{r}
m_int_high_inc = m_int_inc2 %>% estimable(
                   c("(Intercept)" = 0,      # beta0
                     "CP_c"       = 1,      # beta1
                     "income_level_2Higher income" = 0,      # beta2
                     "CP_c:income_level_2Higher income"  = 1),  # beta3
                   conf.int = 0.95)
```

 

**Our conclusion:** For countries with high income, mean life expectancy increases `r round(m_int_high_inc$Estimate, 3)` years for every one additional cell phone per 100 people (95% CI: `r round(m_int_high_inc$Lower.CI, 3)`, `r round(m_int_high_inc$Upper.CI, 3)`).

## If our example had an effect measure modifier

```{r}
m_int_low_inc = m_int_inc2 %>% estimable(
                   c("(Intercept)" = 0,      # beta0
                     "CP_c"       = 0,      # beta1
                     "income_level_2Higher income" = 1,      # beta2
                     "CP_c:income_level_2Higher income"  = 0),  # beta3
                   conf.int = 0.95)
```

- None of our examples had a significant interaction, so it's hard to demonstrate exactly how we would report this

- Let's say, **just for example**, that income had a significant interaction with CP
  - How would we report this to an audience??
  
- Here's how to report on an interaction/EMM:

  - We found that a country's income status (high or low) is a significant effect measure modifier on number of cell phones (*include p-value for interaction test here*). For countries with high income, mean life expectancy increases `r round(m_int_high_inc$Estimate, 3)` years for every additional cell phone per 100 people (95% CI: `r round(m_int_high_inc$Lower.CI, 3)`, `r round(m_int_high_inc$Upper.CI, 3)`). For countries with low income, mean life expectancy increases `r round(m_int_low_inc$Estimate, 3)` years for every additional cell phone per 100 people (95% CI: `r round(m_int_low_inc$Lower.CI, 3)`, `r round(m_int_low_inc$Upper.CI, 3)`).
