---
title: "Homework 1 Answers"
subtitle: "BSTA 512/612"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

Answers are not necessarily complete! This is just meant to serve as a check if you are stuck. 

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 2, fig.width = 4)

library(tidyverse)
library(rstatix)
library(broom)
library(gt)
library(janitor)
library(here)
library(gridExtra)

```

## Directions

This document provides the solutions (final answers) for Homework 1. You may use this to double-check your final numbers and conclusions. This document does not contain the code or derivations required to arrive at the answer.

## Questions

The following questions were adapted from [this textbook](https://catalog.lib.uchicago.edu/vufind/Record/7546272).

### Question 1

#### Part a

::: blue
**Answer:** 0.6306
:::

#### Part b

::: blue
**Answer:** 11.689
:::

### Question 2


#### Part a

::: blue
**Answer:** 
:::

#### Part b


::: blue
**Answer:** .
:::

#### Part c

::: blue
**Answer:** (69.49, 70.51)
:::

### Question 3

::: blue
**Answer:** fail to reject
:::

### Question 4

::: blue
**Answer:** 
:::

### Question 5


::: blue
**Answer:** Any function from `dplyr` not discussed in class (e.g., `arrange`, `slice`, `distinct`, `sample_n`) is acceptable.
:::

### Question 6


#### Part a

::: blue
**Answer:**

```{r}
#| echo: false

library(readxl)
# you might need to update the location of the data file
# you can choose whatever name you like for the tibble when loading it into R's workspace 
ch05q01 <- read_excel(here("data/CH05Q01.xls"))
```

```{r}
#| fig-width: 8
#| fig-height: 4
#| echo: false

names(ch05q01)

G1 <- ggplot(ch05q01, aes(x = AGE, y = DRYWGT)) +
  geom_point() + 
  labs(title = "Scatterplot of Age vs. Dry Weight", 
       x = "Age (days)", 
       y = "Dry weight")

G2 <- ggplot(ch05q01, aes(x = AGE, y = LOGDRYWG)) +
  geom_point() +
  labs(title = "Scatterplot of Age vs. Log Dry Weight", 
       x = "Age (days)", 
       y = "Log dry weight")

library(gridExtra)
grid.arrange(G1, G2, nrow = 1)
```

:::

#### Part b

::: blue
**Answer:**
:::

#### Part c

::: blue
**Answer:**

1.  $\widehat{\text{Dry weight}} = -1.885 + 0.235 \cdot Age$

2.  $\widehat{\log_{10}(\text{Dry weight})} = -2.689 + 0.196 \cdot Age$

OR 

1.  $\widehat{Y} = -1.885 + 0.235 \cdot X$

2.  $\widehat{Z} = -2.689 + 0.196 \cdot X$
:::

#### Part d

::: blue
**Answer:**
:::

#### Part e

::: blue
**Answer:**

* **Slope:** 0.196 (95% CI: 0.185, 0.207)
* **Intercept:** -2.689 (95% CI: -2.809, -2.569)
:::

#### Part f

::: blue
**Answer:**

* **Log-scale:** -1.122
:::

### Question 7

#### Part a

::: blue
**Answer:** `weights`
:::

#### Part b

::: blue
**Answer:**
:::

#### Part c

::: blue
**Answer:** No
:::

### Question 8

::: blue
**Answer:**
:::