---
title: "Homework 2 Answers"
subtitle: "BSTA 512/612"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    self-contained: true
    code-block-bg: "#fff"
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

**Answers are not necessarily complete!** This is just meant to serve as a check if you are stuck. 

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(rstatix)
library(broom)
library(broom.helpers)
library(gt)
library(janitor)
library(readxl)
library(haven)
library(here)
library(gtsummary) ## Added this package!!

```

## Questions

### Question 1

```{r}
#| echo: false
fatal_dep = read_sas(here("data/completedata.sas7bdat"))
```

#### Part a

**Answer:**

::: blue
```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center

ggplot(data = fatal_dep, aes(x = Fatalism, y = Depression)) +
  geom_point() +
  labs(title = "Scatterplot of Depression vs. Fatalism")
```
:::

#### Part b

**Answer:**

::: blue
```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
#| message: false


fd_model = lm(Depression ~ Fatalism, 
              data = fatal_dep)

tidy(fd_model) %>% gt() %>%
  tab_options(table.font.size = 15) %>% fmt_number(decimals = 4)

```
:::

#### Part c

**Answer:**

::: blue
```{r}
fatal_dep2 = fatal_dep %>%
  mutate(FatalismC = Fatalism - median(Fatalism))
```


```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
#| message: false

ggplot(data = fatal_dep2, aes(x = FatalismC, y = Depression)) +
  geom_point() +
  # geom_smooth(method = lm, se = FALSE, color = "#F14124") +
  labs(title = "Scatterplot of Depression vs. Fatalism", 
       x = "Median centered Fatalism")
```
:::

#### Part d

**Answer:**

::: blue
```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
#| message: false


fd_model2 = lm(Depression ~ FatalismC, 
              data = fatal_dep2)

tidy(fd_model2) %>% gt() %>%
  tab_options(table.font.size = 15) %>% fmt_number(decimals = 4)
```

:::

#### Part e

**Answer:**

::: blue
:::

### Question 2

```{r}
# first install the palmerpenguins package
# install.packages("palmerpenguins")
library(palmerpenguins)
data(penguins)

# run the command below to learn more about the variables in the penguins dataset
# ?penguins
```

#### Part a

::: blue
```{r}
flippers_species_ave <- penguins %>% 
  group_by(species) %>% 
  summarise(
    flipper_ave = mean(flipper_length_mm, na.rm = T)
    )

flippers_species_ave 
```
:::

#### Part b

::: blue
```{r}
# Note: In class I used fun.y within stats_summary.
# I noticed that this is now deprecated, and I used fun below instead.
# Both still work.

ggplot(penguins, aes(x = species, 
                   y = flipper_length_mm, 
                   color = species)) +
  geom_jitter(size = 1, alpha = .6, width = 0.2, 
              show.legend = FALSE) +     # removed legend since not needed
  stat_summary(fun = mean, geom = "point", size = 8, shape = 18, 
               show.legend = FALSE) +    # removed legend since not needed
  labs(x = "Species", 
       y = "Flipper lengths (mm)",
       title = "Flipper lengths by penguin species",
       caption = "Diamonds = species averages") 
  
```
:::

#### Part c

Write out the fitted regression equation that models the flipper length by penguin species. Use LaTeX math markup or insert an image of your equation. Do not yet insert values for the regression coefficients, i.e. use the generic coefficients $\widehat{\beta}_0, \widehat{\beta}_1$. Use Adelie as the reference level.

::: blue

:::

#### Part d

::: blue
```{r}
# names(penguins)
model3 <- lm(flipper_length_mm ~ species, data = penguins)
tidy(model3, conf.int=T) %>% gt() # don't need to have confidence intervals, but it will help in next part
```
:::

#### Part e

How do we interpret each of the regression coefficients for this model? *Write out a separate interpretation for each of the coefficients.*

::: blue

:::

#### Part f

Calculate the mean flipper length (and the 95% CI) of penguins in the Chinstrap and Gentoo species using the `predict()` function.

::: blue
```{r}
newdata = data.frame(species = c("Chinstrap", "Gentoo")) 
pred = predict(model3, newdata=newdata, interval="confidence")
```

Chinstrap penguins: `r round(pred[1,1], 2)` mm
:::
