---
title: "Homework 3"
subtitle: "BSTA 512/612"
author: "Your name here!!!"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(rstatix)
library(broom)
library(gt)
library(janitor)
library(readxl)
library(gridExtra)
library(ggfortify)  # autoplot(model)
library(gtsummary)
library(describedata) # gladder()
library(haven)
library(here)
```

## Directions

-   [Download the `.qmd` file here.](https://github.com/nwakim/BSTA_512_26W/blob/main/homework/HW_03.qmd)

-   You will need to download the datasets. Use [this link to download](https://ohsuitg-my.sharepoint.com/:f:/r/personal/wakim_ohsu_edu/Documents/Teaching/Classes/BSTA_512_26W/Student_BSTA_512/Data/Homework?csf=1&web=1&e=es5Kgh) the HW3 datasets needed in this assignment. If you do not want to make changes to the paths set in this document, then make sure the files are stored in a folder named "data" that is housed in the same location as your HW3 `.qmd` file.

-   Please upload your homework to Sakai. **Upload both your `.qmd` code file and the rendered `.html` file**

-   For each question, make sure to include all code and resulting output in the html file to support your answers

-   Show the work of your calculations using R code within a code chunk. Make sure that both your code and output are visible in the rendered html file. This is the default setting.

-   Write all answers in complete sentences as if communicating the results to a collaborator.

::: callout-tip
It is a good idea to try rendering your document from time to time as you go along! Note that rendering automatically saves your `.qmd` file and rendering frequently helps you catch your errors more quickly.
:::

## Questions


### Question 1 {#sec-L06-SLR-More_inf}

This question and data are adapted from [this textbook](https://catalog.lib.uchicago.edu/vufind/Record/7546272).

In an experiment designed to describe the dose–response curve for vitamin K, individual rats were depleted of their vitamin K reserves and then fed dried liver for 4 days at different dosage levels. The response of each rat was measured as the concentration of a clotting agent needed to clot a sample of its blood in 3 minutes. The results of the experiment on 12 rats are given in the following table; values are expressed in common logarithms for both dose and response.

```{r}
clot = read_excel(here("./data/CH05Q09.xls"))
clot %>% gt() %>%
  cols_label(RAT = md("**Rat**"),
             LOGCONC = md("**Log10 Concentration (Y)**"),
             LOGDOSE = md("**Log10 Dose (X)**"))
```

Use the log-transformed values as given in the dataset.

Use the following scatterplot to build your answers off of:

```{r}
#| fig-height: 5
#| fig-width: 6.5
#| fig-align: center
#| echo: false

ggplot(clot, aes(x = LOGDOSE,
                 y = LOGCONC)) +
  geom_point(size = 2) +
  labs(x = "Log10 dose (X)", 
       y = "Log10 concentration (Y)",
       title = "Concentration of clotting agent needed vs. dosage levels of liver") +
    theme(axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        title = element_text(size = 12))
```

#### Part a

Fit a linear regression model to the data and add the regression line to the plot.

#### Part b

Use R to create the ANOVA table for the regression described in the exercise.

#### Part c

Using the F-test, determine whether there is an association between the log10 concentration and log10 dose.

::: callout-note
Make sure to include all needed steps for an F-test. Calculating the F test statistic (step 5) is not needed if you use the ANOVA table. Make sure your conclusion connects back the research context.
:::

#### Part d

Rewrite your hypothesis test in Part c to show the null and alternative models that we are testing. Did we reject the smaller (reduced) model?

::: callout-note
You do not need to go through the hypothesis test process again. A quick statement on rejecting or not is okay.
:::

::: callout-tip
If you prefer to write out the models by hand, remember that you can take a picture of your work and insert it into this document. HW0 can be a good reference for how we've done this before.
:::

### Question 2 {#sec-L06-SLR-More_inf}

We will continue to work with the study and dataset from Question 2 above.

#### Part a

Find the correlation coefficient between the two variables. Is the value consistent with your description of the relationship in Question 2? Why or why not?

#### Part b

Calculate the coefficient of determination using linear regression summary output. Can we also calculate the coefficient of determination from the ANOVA in Question 2?

#### Part c

Give an interpretation of the coefficient of determination in the context of the study.


### Question 3 {#sec-L07-L08-Assumptions-Diagnostics}

A high respiratory rate is a potential diagnostic indicator of respiratory infection in children. To judge whether a respiratory rate is “high” however, a physician must have a clear picture of the distribution of normal rates. To this end, Italian researchers measured the respiratory rates (in breaths/minute) of 618 children between the ages of 15 days and 3 years (measured in months).

The data and problem framing came from the `Sleuth3` package. Please make sure to run the following code to load the data. You can directly access the dataset `ex0824` from the package. I have included a new assignment of the data to `q1_data` if you would like to use that.

```{r}
#| message: false

if(!require(Sleuth3)) { install.packages("Sleuth3"); library(Sleuth3) }
q1_data = ex0824 
```

#### Part a

Create a scatterplot of the dependent and independent variables with both the best-fit line and a smoothed curve through the points. Describe the relationship between the dependent and independent variables, and also comment on whether you think it is reasonable to use a linear regression to model the relationship.

#### Part b

Write out the population regression model for the simple linear regression model. Please leave the variables untransformed for now.

#### Part c

Fit the regression model, display the regression table, and write out the fitted regression line.

#### Part d

Assess the normality of the model's fitted residuals by creating a histogram, density plot, and boxplot of the residuals to visually inspect the distribution of the residuals, and describe any deviations from normality.


#### Part e

Assess the normality of the model's fitted residuals by creating QQ plot of the residuals.

Bonus work, but not required: Compare the QQ plot to 4 such plots simulated from normal data, and discuss why or why not the residuals could have come from a normal distribution.

#### Part f

Test the normality of the model's fitted residuals and comment on whether the test's conclusion is consistent with your visual inspection or not. Make sure to include the hypotheses, needed R code, and a conclusion to the test based on the p-value ([as shown in these slides](https://nwakim.github.io/W2024_BSTA_512/slides/06_SLR_Diagnostics.html#/n-shapiro-wilk-test-of-normality)).

#### Part g

Create a residual plot using ggplot and the residuals. Discuss what this means in the context of our model assumptions.

#### Part h

Determine whether there are any observations with high leverage. Please use the cutoff, $h_i > 6/n$. If there are observations with high leverage, print the observations and state how many high leverage points there are.

#### Part i

Print the 10 observations with highest Cook's distance. If there are observations with high Cook's distance ($d_i >1$), state how many observations have high Cook's distance.

#### Part j

Create a histogram for rate. Describe its distribution shapes.

#### Part k

Using the above histogram, and Tukey's ladder of transformations, discuss the range of transformations that will be appropriate for **Rate**. Explain your reasoning.

Then use `gladder()` to decide on two possible transformations. Explain your reasoning.

Note: questions below will ask about model fit with the transformations. For now, just explain why you chose the two that you did.

#### Part l

Add the two rate transformations you chose above to the dataset. You do not need to print any output, just make sure the code is visible.

#### Part m

Create scatterplots using two transformed rates and age. Discuss if either transformation potentially improves the model fit. Explain why or why not. Note: including lines will help!

#### Part n

Using one of the transoformed outcomes, fit the regression model, display the regression table, and write out the fitted regression line.

#### Part o

Assess the normality of the model’s fitted residuals by creating QQ plot of the residuals. Does the transformation improve the QQ plot?

#### Part p

Create a residual plot using ggplot and the residuals. Discuss what this means in the context of our model assumptions. Does the transformation improve our model assumptions?

#### Part q

Between the model with the untransformed outcome and the transformed outcome, which would you recommend using for analysis? (Hint: there are pros and cons to both models)

