---
title: "Lab 2: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#213c96"
date: "`r library(here); source(here('class_dates.R')); w6d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 2: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)
```


## Notes from me and the TAs

- Make sure your .Rproj is open! 
- Please use the template .qmd that I give you!
- Univariate exploration should include all the race variables (indicators for each race)
- You should not have any NA's
  - Please make sure you coded EACH variable correctly...
  - Saw people missing a gender identity
- Please interpret all the coefficient estimates!!
  - Everyone should have more than one coefficient
  - Good reference for self-perception: "Neither ...."
- The whole point of plotting each variable is to start understanding trends in the data...
  - Looking at some plots, I don't think you can do that...
  
## Notes on plots

- Make sure your categorical variable is presented in order
- Make sure your plots are clear and easy to read (e.g., axis labels, title, etc.)
    - If the category names are too long, tilt the names or make new line
- Histograms are for CONTINUOUS variables, bar plots are for CATEGORICAL variables
- Do NOT use the continuous variables for your categorical predictor!!
  - I'm seeing a lot of people use `important_001` instead of their categorical/factor version of this variable
- Univariate vs bivariate
- Colors in histograms are fun, but not necessary!! 
    - Do not use a different color for each plot - color is meant to convey information! 
    - The main thing it does it alert me that you might be using AI bc AI does this a lot!

## Get into groups of 2-4

- No more than 4! 
- Try to group with people with whom you do not already work on labs 
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(2)
```

## Introduce Research Question

- Share your research question with each other! 
  - Should be using the same outcome + your main variable
- Share your other variables

 

```{r}
#| echo: false
countdown::countdown(3)
```

## Share your bivariate plots

- Can you get an idea of the relationship between your main variable and the outcome from the plot?
- Is your predictor categorical?
- Do you see any trends?

```{r}
#| echo: false
countdown::countdown(10)
```


## Simple logistic regression

- Does everyone have the correct interpretations of the coefficient estimates?
- What's your reference group? Does this make sense?
- What trends are you seeing? 

```{r}
#| echo: false
countdown::countdown(10)
```

