---
title: "Poster Help Session"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#213c96"
date: "`r library(here); source(here('class_dates.R')); w10d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Poster info"
    highlight-style: ayu
    html-math-method: mathjax
execute:
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false

library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(broom)
library(rstatix)
library(gt)
library(readxl)
library(describedata) # gladder()
library(gridExtra)   # grid.arrange()
library(ggfortify)  # autoplot(model)
library(gtsummary)
library(plotly) # for plot_ly() command
library(GGally) # for ggpairs() command 
library(ggiraphExtra)   # for ggPredict() command
library(dvmisc)
library(expss)

load(here("data", "gapminder2022.RData"))
gapm = gapm %>%
  mutate(fs_order = case_when(freedom_status == "F" ~ 3,
                                freedom_status == "PF" ~ 2,
                                freedom_status == "NF" ~ 1), 
         freedom_status = case_match(freedom_status, 
                                     "NF" ~ "Not free", 
                                     "PF" ~ "Partly free", 
                                     "F"  ~ "Free"),
         freedom_status = fct_relevel(freedom_status, "Not free", "Partly free", "Free")) %>%
  mutate(income_level_4 = fct_relevel(income_level_4, 
                                      "Low income", 
                                      "Lower middle income",
                                      "Upper middle income",
                                      "High income")) %>%
  mutate(income_level_2 = case_when(income_level_4 %in% c("Low income", "Lower middle income") ~ "Lower income",
                                      income_level_4 %in% c("Upper middle income", "High income") ~ "Higher income")) %>%
  mutate(income_level_2 = fct_relevel(income_level_2, "Lower income", "Higher income")) %>%
  mutate(CP_c = cell_phones_100 - mean(cell_phones_100, na.rm=T)) %>%
  mutate(co2_emissions = co2_emissions/1000/1000)

gapm2 = gapm %>% select(cell_phones_100, life_exp, vax_rate, freedom_status, income_level_4, basic_sani,  co2_emissions, happiness_score)
gapm2 = gapm2 %>% 
  mutate(BS_q = quant_groups(basic_sani, groups = 4) %>% factor(labels = c("Quartile 1", "Quartile 2", "Quartile 3", "Quartile 4")),  
         BS_q = fct_relevel(BS_q, "Quartile 1", "Quartile 2", "Quartile 3", "Quartile 4"))

#load(here("./lessons/15_MLR_Diagnostics/final_mod.rda"))
final_model = gapm2 %>% lm(formula = life_exp ~ cell_phones_100 + BS_q + freedom_status + income_level_4 + vax_rate + happiness_score)

theme_set(theme_bw(base_size = 22))
```

## Some lab 4 notes

- I added some code to the instructions to order the labels in your forest plot!
- For interactions, the only possible interaction you should include should be your subquestion!!
  - Multiple interaction is a little too complicated right now
- If you have an interaction, you should not include it in your forest plot (more to come)
  - Best to include the multivariable visualization you made in Lab 3
  - Describe how the effect modifier has changed the effect of your main variable
  - Make sure you are interpreting your coefficients correctly if you have an interaction!!
  
## Get into groups of 2-4

- No more than 4! 
- Introduce yourself and your research question
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(2)
```

## Purposeful Selection

- Did everyone follow the steps correctly? Please review the decision rules among yourselves
- Parts that seemed especially difficult
  - Change in coefficient estimate
  
```{r}
#| echo: false
countdown::countdown(8)
```  

## Some notes

-   I will NOT be counting exact bullet points
    -   If it says 2-3, anything less than 4 is okay
    -   If it says 5-6, it shouldn't be 1 and it shouldn't be \>10
-   Do NOT make bullet points really long!
-   I tried the QMD poster, and it is REALLY hard to troubleshoot
    -   I highly suggest making your poster in powerpoint!

```{css}
.reveal .sourceCode {
  font-size: 0.9em !important;
}

```

## Background

-   **Length: 5-8 bullets**
-   **Purpose:** Introduce the research question and why it is important to study
-   This section is non-technical.
    -   By reading just the introduction and conclusion, someone without a technical background should have an idea of what they study was about, why it is important, and what the main results are
-   You may start with your bullets from Lab 1, but you should edit it and make sure it flows into your report well!
-   Should contain some references

## Methods

-   **Length: 8-10 bullets**
-   **Purpose:** Describe the analyses that were conducted and methods used to select variables and check diagnostics
-   **Some important methods to discuss** (You may divide these into your sections, not necessarily with these names)
    -   General approach to the dataset
    -   Variables and variable creation
    -   Model building: we performed purposeful selection
    -   Final model
    -   Model diagnostics
    
## Methods for life expectancy {.smaller}

- Data were collected from Gapminder and World Bank with 197 countries in 2011
- We performed a complete case analysis on 105 countries
- We generated a categorized variable for basic sanitation and income level
  - Basic sanitation used quartiles
  - Income levels used the specified groupings by Gapminder
- We used purposeful model selection, a combination of field expertise and statistical methods, to determine the final model
- We performed linear regression on our outcome, life expectancy, with a main effect for cell phones per 100 people while adjusting for confounders $$\widehat{\text{LE}} = \widehat{\beta}_0 + \widehat{\beta}_1 \text{FLR} + \text{other confounders}$$
  - We adjusted for: income levels, basic sanitation, vaccination rate, freedom status, and happiness score
- We investigated model assumptions and diagnostics using standardized residuals, leverage, Cook's distance, and variance inflation factors (VIFs)
- We used R version 4.5.1 to analyze data
  

## Results: What should be in your results?

1.  Table 1 (table summary in Lesson 2)

2.  Regression table or Forest plot

3.  One additional figure or table to help understand your question

4.  Interpretation(s) of important coefficient estimates 

## 1. Table 1 (table summary in Lesson 2)

::::: columns
::: {.column width="60%"}
```{r}
#| echo: true
#| eval: false

library(gtsummary)

gapm2_vars = gapm2 %>% 
  select(cell_phones_100, 
         vax_rate, 
         freedom_status, 
         income_level_4, 
         basic_sani,
         happiness_score)

tbl_summary(
  gapm2_vars, 
  label = list(
    cell_phones_100 ~ "Cell phones per 100 people", 
    basic_sani ~ "Basic sanitation (%)",
    freedom_status ~ "Freedom status",
    income_level_4 ~ "Income level",
    vax_rate ~ "Vaccination rate (%)",
    happiness_score ~ "Happiness score"
    ), 
  statistic = list(all_continuous() ~ "{mean} ({sd})")
  )
```
:::

::: {.column width="40%"}
```{r}
library(gtsummary)

gapm2_vars = gapm2 %>% 
  select(cell_phones_100, 
         vax_rate, 
         freedom_status, 
         income_level_4, 
         basic_sani,
         happiness_score)

tbl_summary(
  gapm2_vars, 
  label = list(
    cell_phones_100 ~ "Cell phones per 100 people", 
    basic_sani ~ "Basic sanitation (%)",
    freedom_status ~ "Freedom status",
    income_level_4 ~ "Income level",
    vax_rate ~ "Vaccination rate (%)",
    happiness_score ~ "Happiness score"
    ), 
  statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>% 
  as_gt() %>% 
  tab_options(table.font.size = 26)
```
:::
:::::

## 2. Regression table or Forest plot

::::: columns
::: {.column width="52%"}
```{r}
#| echo: true
#| eval: false

tbl_regression(
  final_model, 
  label = list(
    cell_phones_100 ~ "Cell phones per 100 people", 
    BS_q ~ "Basic sanitation (%)",
    freedom_status ~ "Freedom status",
    income_level_4 ~ "Income level",
    vax_rate ~ "Vaccination rate (%)",
    happiness_score ~ "Happiness score"
    )) %>% 
  as_gt() %>% 
  tab_options(table.font.size = 25) %>%
  cols_width(label ~ px(250))
```
:::

::: {.column width="48%"}
```{r}
tbl_regression(
  final_model, 
  label = list(
    cell_phones_100 ~ "Cell phones per 100 people", 
    BS_q ~ "Basic sanitation (%)",
    freedom_status ~ "Freedom status",
    income_level_4 ~ "Income level",
    vax_rate ~ "Vaccination rate (%)",
    happiness_score ~ "Happiness score"
    )) %>% 
  as_gt() %>% 
  tab_options(table.font.size = 25) %>%
  cols_width(label ~ px(250))
```
:::
:::::

## 2. Regression table or Forest plot

::::: columns
::: {.column width="50%"}

- This is a fun one to investigate!
- Stick to the regression table if you are having trouble with this!
```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-summary: "Code for forest plot"

library(broom.helpers)

model_tidy = tidy_and_attach(final_model, conf.int=T) %>%
  tidy_remove_intercept() %>%
  tidy_add_reference_rows() %>% tidy_add_estimate_to_reference_rows() %>%
  tidy_add_term_labels() %>%
  mutate(label = fct_rev(fct_inorder(label))) %>%
  mutate(var_label = case_match(var_label, 
                               "BS_q" ~ "Basic sanitation",
                               "freedom_status" ~ "Freedom status",
                               "income_level_4" ~ "Income level",
                               "vax_rate" ~ " ",
                               "happiness_score" ~ "  ",
                               "cell_phones_100" ~ "    ",
                               .default = var_label
                         )) %>%
  mutate(label = case_match(label, 
                               "vax_rate" ~ "Vaccination rate (%)",
                               "happiness_score" ~ "Happiness score",
                               "cell_phones_100" ~ "Cell phones per 100 people",
                               .default = label
                         ))

ggplot(data=model_tidy, aes(y=label, x=estimate, xmin=conf.low, xmax=conf.high)) + 
  facet_grid(rows = vars(var_label), scales = "free",
             space='free_y', switch = "y") + 
  geom_point(size = 3) +  geom_errorbarh(height=.2) + 
  geom_vline(xintercept=0, color='#C2352F', linetype='dashed', alpha=1) +
  theme_classic() +
  labs(x = "Beta", y = "Variables") +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 16), 
        title = element_text(size = 16), strip.placement = "outside", 
        strip.text.y.left = element_text(size = 16, angle = 0), 
        strip.background = element_blank())
```
:::

::: {.column width="50%"}
```{r}
#| fig-height: 10
#| fig-width: 9
#| fig-align: center

library(broom.helpers)

model_tidy = tidy_and_attach(final_model, conf.int=T) %>%
  tidy_remove_intercept() %>%
  tidy_add_reference_rows() %>% tidy_add_estimate_to_reference_rows() %>%
  tidy_add_term_labels() %>%
  mutate(label = fct_rev(fct_inorder(label))) %>%
  mutate(var_label = case_match(var_label, 
                               "BS_q" ~ "Basic \n sanitation",
                               "freedom_status" ~ "Freedom \n status",
                               "income_level_4" ~ "Income \n level",
                               "vax_rate" ~ " ",
                               "happiness_score" ~ "  ",
                               "cell_phones_100" ~ "    ",
                               .default = var_label
                         )) %>%
  mutate(label = case_match(label, 
                               "vax_rate" ~ "Vaccination rate (%)",
                               "happiness_score" ~ "Happiness score",
                               "cell_phones_100" ~ "Cell phones per \n 100 people ",
                               .default = label
                         ))


ggplot(data=model_tidy, aes(y=label, x=estimate, xmin=conf.low, xmax=conf.high)) + 
  facet_grid(rows = vars(var_label), scales = "free",
             space='free_y', switch = "y") + 
  geom_point(size = 3) +  geom_errorbarh(height=.2) + 
  geom_vline(xintercept=0, color='#C2352F', linetype='dashed', alpha=1) +
  theme_classic() +
  labs(x = "Beta", y = "Variables") +
  theme(axis.title = element_text(size = 16), axis.text = element_text(size = 16), 
        title = element_text(size = 16), strip.placement = "outside", 
        strip.text.y.left = element_text(size = 16, angle = 0), 
        strip.background = element_blank())
```
:::
:::::

## Reporting results

-   **Length: 5-8 bullets**
-   **Purpose:** Interpret the results of your analyses and explain what they mean in the context

 

-   Let's just say basic sanitation was my main variable
-   In my results, I would say something like:
    - In general, countries in higher quartiles of basic sanitation have higher life expectancy adjusting for all other variables in the final model
    - For example, based on the estimated coefficients in the forest plot:
      - Countries in the 4th quartile of basic sanitation have an expected life expectancy that is `r round(model_tidy$estimate[5], 2)` years higher than countries in the 1st quartile of basic sanitation (95% CI: `r round(model_tidy$conf.low[5], 2)`, `r round(model_tidy$conf.high[5], 2)`).
      - Similar interpretations can be made for coefficients of the 2nd and 3rd quartiles of basic sanitation
    



## Important note if you have interactions

-   There is extra care to reporting results from interactions

-   Please make sure you are interpreting your coefficients correctly!!

-  If you have an interaction, you cannot interpret the main effects without considering the interaction term
  - You can exclude the interaction coefficients from your forest plot

## Example of main effects interpretation with an interaction
- For example, if we have an interaction between importance of weight and political ID, our model might look like: $$\begin{aligned}
\widehat{\text{IAT}} = &\widehat{\beta}_0 + \widehat{\beta}_1 I(\text{imp} = \text{Slightly}) + \widehat{\beta}_2 I(\text{imp} = \text{Moderately}) + \widehat{\beta}_3 I(\text{imp} = \text{Very}) + \\
&\widehat{\beta}_4 I(\text{imp} = \text{Extremely}) + \widehat{\beta}_5 I(\text{pol} = \text{Strongly liberal}) + \widehat{\beta}_6 I(\text{pol} = \text{Moderately liberal}) + \\
& \widehat{\beta}_7 I(\text{pol} = \text{Slightly liberal}) + \widehat{\beta}_8 I(\text{pol} = \text{Slightly conservative}) + \\
& \widehat{\beta}_9 I(\text{pol} = \text{Moderately conservative}) + \widehat{\beta}_{10} I(\text{pol} = \text{Strongly conservative}) + \\
& ... \text{confounders} ... + \\
& \widehat{\beta}_{11} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Strongly liberal}) + \\ 
&\widehat{\beta}_{12} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Moderately liberal}) + \\
& \widehat{\beta}_{13} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Slightly liberal}) + \\
&\widehat{\beta}_{14} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Slightly conservative}) + \\
& \widehat{\beta}_{15} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Moderately conservative}) + \\
& \widehat{\beta}_{16} I(\text{imp} = \text{Slightly}) \cdot I(\text{pol} = \text{Strongly conservative}) \\
\end{aligned}$$
