---
title: "Lab 1: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#213c96"
date: "`r library(here); source(here('class_dates.R')); w3d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 1: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Notes from me and the TAs



## Get into groups of 2-4

- No more than 4! 
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Introduce Research Question

- Share your research question with each other! 
- Share your other variables
- Is everyone using the appropriate variables?
- Did anyone see any potential issues with how many individuals are in each category of a variable?

```{r}
#| echo: false
countdown::countdown(15)
```

## Double check code and saved data

- Did everyone in the group successfully save the dataset?
- Can you open it right now?

```{r}
#| echo: false
countdown::countdown(5)
```

## Compiled introductions

- What major points did you include in your introduction?
- Would you change anything after discussing with your group?

```{r}
#| echo: false
countdown::countdown(5)
```