---
title: "Lab 3: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "05/12/2025"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 3: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  eval: false
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Lab 3 Feedback

- Some fo you did main covariate + 4 other variables and some did main covariate + 5 other variables
  - Okay either way
- In presentation of odds ratios (forest plot and tables), it's important to include the reference group
  - Gives the reader context for the comparison


## Get into groups of 2-4

- No more than 4! 
- Introduce yourself and your research question
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Help with forest plot: ordering categories

```{r}
source("/Users/wakim/Library/CloudStorage/OneDrive-OregonHealth&ScienceUniversity/Projects/Teaching_DEIJ_Datasets/Datasets/Food_insecurity/Code/Labs/Lab_02_work.R")
```

Let's say I ran this model:
```{r}
model = wbns4 %>% 
  glm(formula = FOOD_INSEC ~ PPINCIMP + XACSNET + FAMSIZE, family = binomial)
```

## Help with forest plot: ordering categories {.smaller}

```{r}
library(broom.helpers)
MLR_tidy0 = tidy_and_attach(model,  conf.int=T, exponentiate = T) %>%
  tidy_remove_intercept() %>%
  tidy_add_reference_rows() %>%
  tidy_add_estimate_to_reference_rows() %>%
  tidy_add_term_labels()

glimpse(MLR_tidy0)

```

## Help with forest plot: ordering categories {.smaller}
```{r}

plot_MLR0 = ggplot(data=MLR_tidy0, 
       aes(y=label, x=estimate, xmin=conf.low, xmax=conf.high)) + 
  geom_point(size = 3) +  geom_errorbarh(height=.2) + 
  
  geom_vline(xintercept=1, color='#C2352F', linetype='dashed', alpha=1) +
  theme_classic() +
  
  facet_grid(rows = vars(var_label), scales = "free",
             space='free_y', switch = "y") + 
  
  labs(x = "OR (95% CI)", 
       title = "Odds ratios of food insecurity") +
  theme(axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        title = element_text(size = 12), 
        axis.title.y=element_blank(), 
        strip.text = element_text(size = 12), 
        strip.placement = "outside", 
        strip.background = element_blank())
```

## Help with forest plot: ordering categories {.smaller}

```{r}
#| fig-width: 5
#| fig-height: 7

plot_MLR0
```

## Help with forest plot: ordering categories {.smaller}
```{r}
MLR_tidy = MLR_tidy0  %>%
              mutate(order = case_match(label, 
                            "Less than $5,000" ~ 1, 
                            "$5,000 to $7,499" ~ 2, 
                            "$7,500 to $9,999" ~ 3, 
                            "$10,000 to $12,499" ~ 4, 
                            "$12,500 to $14,999" ~ 5, 
                            "$15,000 to $19,999" ~ 6, 
                            "$20,000 to $24,999" ~ 7, 
                            "$25,000 to $29,999" ~ 8, 
                            "$30,000 to $34,999" ~ 9, 
                            "$35,000 to $39,999" ~ 10, 
                            "$40,000 to $49,999" ~ 11, 
                            "$50,000 to $59,999" ~ 12, 
                            "$60,000 to $74,999" ~ 13, 
                            "$75,000 to $84,999" ~ 14, 
                            "$85,000 to $99,999" ~ 15, 
                            "$100,000 to $124,999" ~ 16, 
                            "$125,000 to $149,999" ~ 17, 
                            "$150,000 to $174,999" ~ 18, 
                            "$175,000 to $199,999" ~ 19, 
                            "$200,000 to $249,999" ~ 20, 
                            "$250,000 or more" ~ 21, 
                            "1" ~ 1, 
                            "2" ~ 2, 
                            "3" ~ 3, 
                            "4" ~ 4, 
                            "5" ~ 5, 
                            "6" ~ 6, 
                            "7" ~ 7, 
                            "8 or more" ~ 8,
                            .default = 0)) %>%
            mutate(label = fct_reorder(label, desc(order)))
```

## Help with forest plot: ordering categories {.smaller}
```{r}

plot_MLR = ggplot(data=MLR_tidy, 
       aes(y=label, x=estimate, xmin=conf.low, xmax=conf.high)) + 
  geom_point(size = 3) +  geom_errorbarh(height=.2) + 
  
  geom_vline(xintercept=1, color='#C2352F', linetype='dashed', alpha=1) +
  theme_classic() +
  
  facet_grid(rows = vars(var_label), scales = "free",
             space='free_y', switch = "y") + 
  
  labs(x = "OR (95% CI)", 
       title = "Odds ratios of food insecurity") +
  theme(axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        title = element_text(size = 12), 
        axis.title.y=element_blank(), 
        strip.text = element_text(size = 12), 
        strip.placement = "outside", 
        strip.background = element_blank())
```

## Help with forest plot: ordering categories {.smaller}

```{r}
#| fig-width: 5
#| fig-height: 7

plot_MLR
```



## Questions to consider

- How did you handle any categorical variables?
  - Income, family size, number of children, etc
- Are the categories for the forest plot in order?
- How did you interpret your estimated odds ratios?
  - If you have many categories in your main variable, do you see a trend? 
  - What might be a better way to report a trend?
- If you had interactions, how would your interpretation change?
