---
title: "Lab 3: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#213c96"
date: "`r library(here); source(here('class_dates.R')); w8d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 3: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```


## Get into groups of 2-4

- No more than 4! 
- Introduce yourself and your research question
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(2)
```

## Discuss confounders and potential interactions

- This will drive a lot of our model selection


```{r}
#| echo: false
countdown::countdown(10)
```

## Multivariate plots

- Does your plot show all the categories together? 
- AKA can you get a sense of possible confounder/interaction easily from your plot?
  - Can someone unfamiliar with the data get a sense of it in a minute or less?

```{r}
#| echo: false
countdown::countdown(5)
```

## Multiple Linear Regression

- Does everyone have an interpretation for each coefficient for the main variable?
- How would you synthesize the results in a sentence or two?
  - In the poster, it will be good to discuss the trend, then give an example of how you would interpret each coefficient

```{r}
#| echo: false
countdown::countdown(10)
```

